\newglossaryentry{WAD}
{
  name=WAD,
  description={Default format of package files for the DOOM / DOOM II video-games. "WAD" is an acronym for "Where's all the Data?" \cite{wadmeaning} },  plural=WADs
}
\newglossaryentry{lump}
{
	name=Lump,
	description={ Any section of data within a \gls{WAD} file. }
}
\newglossaryentry{thing}
{
	name=Thing,
	description={ Any deployable asset of a DOOM level, such as Enemies, Power-Ups, Weapons, Decorations, Spawners, Etc. }
	plural=Things
}
\newglossaryentry{sector}
{
	name=Sector,
	description={A Sector in a DOOM level is any closed area (with possibly invisible walls) that has a constant floor and ceiling height and texture.  }
	plural=Things
}
\newglossaryentry{deathmatch}
{
	name=Deathmatch,
	description={ Game mode in which two or more players compete against each other in achieving the highest number of kills before a timeout or a player reaches a predetermined score. }
}

\newglossaryentry{MU}{
	name=MU,
	description={Map Units, Coordinate unit used in Doom Rendering Engine. }
}

\newglossaryentry{DU}{
	name=DU,
	description={Doom Units, see \gls{MU} }
}

\newglossaryentry{linedef}{
	name=Linedef,
	description={ Line that connects two vertices in the WAD file specification }
}

\newglossaryentry{sidedef}{
	name=Sidedef,
	description={ A "vertical plane" in the WAD file specification.}
}

\newglossaryentry{featuremap}{
	name=Feature Map,
	description={Image describing a particular aspect of a level},
	plural={Feature Maps}
}

\newglossaryentry{floormap}{
	name=FloorMap,
	description={Image describing which parts of the map are occupied by a floor and which are empty.}
}
\newglossaryentry{heightmap}{
	name=HeightMap,
	description={Image describing the floor height of a level}
}
\newglossaryentry{roommap}{
	name=RoomMap,
	description={Image describing the room segmentation of the level}
}
\newglossaryentry{thingsmap}{
	name=ThingsMap,
	description={Image describing how the game objects are placed inside the level}
}
\newglossaryentry{triggermap}{
	name=TriggerMap,
	description={Image describing doors, lifts and switches and their correlation}
}
\newglossaryentry{wallmap}{
	name=WallMap,
	description={Image describing the impassable walls in a level}
}
\newglossaryentry{floor}{
	name=floor,
	description={Unconnected piece of level, reachable only by means of a teleporter}
}

\newglossaryentry{gan}{
	name=GAN,
	description={Generative Adversarial Network}
}




